Covariate analysis of UFL-Mayo-ISB temporal cortex arrays

### 2016-01-26 code/Rmd/Mayo_TemporalCortex_Array.Rmd
Added differential expression to Mayo temporal cortex data

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

@@

# Query biomart
Illumina2HGNC <- getBM(attributes = c("illumina_humanht_12_v4", "hgnc_symbol"),
                       filters = "illumina_humanht_12_v4", 
                       values = RESIDUAL.GENE_EXPRESSION$illumina_humanht_12_v4,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(Illumina2HGNC, RESIDUAL.GENE_EXPRESSION) %>%
  dplyr::select(illumina_humanht_12_v4, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Temporal Cortex Microarray', parentId = parentId)
CODE <- synStore(CODE)

@@

# Store expression values
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'Voom Adjusted Residual Expression (Dxn added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |

>>>>

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

@@

# Query biomart
Illumina2HGNC <- getBM(attributes = c("illumina_humanht_12_v4", "hgnc_symbol"),
                       filters = "illumina_humanht_12_v4", 
                       values = RESIDUAL.GENE_EXPRESSION$illumina_humanht_12_v4,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(Illumina2HGNC, RESIDUAL.GENE_EXPRESSION) %>%
  dplyr::select(illumina_humanht_12_v4, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Get design matrix
DM = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F], Intercept = F)

# Linear model with new design
FIT = lmFit(EXPR, design=DM$design)

# Fit contrast
contrast = makeContrasts(contrasts="Dxn0-Dxn1", levels = colnames(FIT$coefficients))
FIT = contrasts.fit(FIT, contrasts=contrast)
FIT = eBayes(FIT)

# Get differnetial expression
DE = topTable(FIT, coef=1, number = dim(EXPR)[1])
DE = rownameToFirstColumn(DE, 'illumina_humanht_12_v4') %>%
  left_join(Illumina2HGNC)
p <- ggplot(DE, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p

tmp = DE %>% dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(hgnc_symbol) %>%
  unlist %>% unique
kable(data.frame(Comparison = "Control_vs_AD",
           nDiffExp = length(tmp),
           DiffExpGenes = paste(tmp, collapse = ',')))  
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Temporal Cortex Microarray', parentId = parentId)
CODE <- synStore(CODE)

@@

# Store expression values
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'Voom Adjusted Residual Expression (Dxn added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

# Store differential expression results
write.table(DE, file = 'diffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('diffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |
|  Differential Expression  |  `r DEXP_OBJ$properties$id`      |
